---
title: "Safe expression compilation and exploration"
author: "MultiOmicsDataCompile authors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Safe expression compilation and exploration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Purpose

`MultiOmicsDataCompile` keeps RNA-seq counts and microarray expression in
separate per-dataset experiments. A separate `ExplorationSE` object contains
within-dataset feature z-scores for qualitative visualization. It must not be
used for differential-expression inference or absolute abundance comparisons
between platforms.

## Validate a dataset manifest

```{r manifest}
library(MultiOmicsDataCompile)

manifest <- data.frame(
  ID = c("RNA_EXAMPLE", "ARRAY_EXAMPLE"),
  Technology = c("RNA-Seq", "microarray"),
  Species = c("Homo sapiens", "Human"),
  GEO2R = c(FALSE, FALSE),
  stringsAsFactors = FALSE
)
manifest <- ValidateDatasetManifest(manifest)
manifest[, c("ID", "Technology", "Species")]
```

## Compile expression objects

This small example uses six features. Production workflows normally obtain the
compiled table from `RawDataCompile()`.

```{r compilation}
samples <- c("RNA_A", "RNA_B", "RNA_C", "RNA_D", "ARRAY_A", "ARRAY_B")
compiled <- data.frame(
  FEATURE_ID = paste0("Human:", seq_len(6)),
  SPECIES = "Human",
  ENTREZID = as.character(seq_len(6)),
  SYMBOL = paste0("GENE", seq_len(6)),
  RNA_A = c(10, 20, 30, 40, 50, 60),
  RNA_B = c(12, 18, 35, 44, 48, 61),
  RNA_C = c(30, 40, 20, 55, 70, 80),
  RNA_D = c(28, 44, 19, 51, 72, 84),
  ARRAY_A = c(5.1, 6.2, 7.3, 8.4, 9.5, 10.6),
  ARRAY_B = c(5.4, 6.0, 7.8, 8.1, 9.8, 10.2),
  check.names = FALSE
)

sample_data <- data.frame(
  dataset = c(rep("RNA_EXAMPLE", 4), rep("ARRAY_EXAMPLE", 2)),
  assay.type = c(rep("RNA-Seq", 4), rep("Array", 2)),
  rawcolumnnames = samples,
  condition = c("control", "control", "treated", "treated", "A", "B"),
  row.names = samples,
  stringsAsFactors = FALSE
)

collection <- GenerateRawSE(sample_data, compiled, manifest)
names(collection$RNAseqSE)
names(collection$ArraySE)
SummarizedExperiment::assayNames(collection$RNAseqSE$RNA_EXAMPLE)
SummarizedExperiment::assayNames(collection$ExplorationSE)
```

RNA-seq conditions are always explicit. If any are missing, compilation stops
instead of guessing group membership.

## Where conditions come from

Public repositories do not record a usable condition, so the batch workflow
derives one with `SampleConditions()`. Each manifest comparison code places one
character per sample of the stored raw matrix, making the assignment a lookup
rather than a guess.

```{r conditions}
directory <- tempfile()
dir.create(directory)
saveRDS(
  data.frame(
    ENTREZID = c("1", "2"), SYMBOL = c("A", "B"),
    RNA_A = c(5, 3), RNA_B = c(6, 4), RNA_C = c(9, 1), RNA_D = c(8, 2)
  ),
  file.path(directory, "RNA_Raw.rds")
)

SampleConditions(
  DBPath = directory,
  overview = data.frame(
    ID = "RNA",
    ComparisonVector = "0011",
    FCColumnNames = "NASH-Ctrl_logFC,NASH-Ctrl_Pvalue,NASH-Ctrl_AdjPValue",
    stringsAsFactors = FALSE
  )
)
```

Pass the result to `MetDataCompile(conditions = )` so that the compiled sample
metadata carry a `condition` column. Samples that every comparison excludes are
labelled `"Unassigned"` rather than assigned to a group.

## RNA-seq quality summaries

```{r qc, fig.width=7, fig.height=4}
qc <- RNAseqQC(collection$RNAseqSE$RNA_EXAMPLE)
qc$sample_metrics[, c("sample", "library_size", "detected_features")]
qc$pca_plot
```

## Reproducible batch execution

`makeDirectory()` copies a `_targets.R` template into the processing directory.
Set `MULTIOMICS_DATA_DIR` and run `targets::tar_make()` from its `Scripts`
directory. Expensive downloads and processing stages are then rerun only when
their inputs or code change.

## Session information

```{r session-info}
sessionInfo()
```
